---
title: "Machine Learning Prediction Report"
author: "Widiani"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

This project aims to predict exercise performance based on sensor data collected from wearable devices. The dataset includes movement data measured from the belt, forearm, arm, and dumbbell of participants. The primary goal is to classify exercises correctly using machine learning techniques.

---

## Data Loading & Cleaning

The dataset is read from the current directory, and preprocessing includes removing high missing value columns and irrelevant metadata.

```{r}
library(caret)
library(randomForest)
library(ggplot2)
library(dplyr)
library(readr)

# Define dataset paths
train_path <- "pml-training.csv"
test_path <- "pml-testing.csv"

# Load dataset
training <- read_csv(train_path)
testing <- read_csv(test_path)

# Remove columns with excessive missing values
na_threshold <- 0.95
na_counts <- colSums(is.na(training)) / nrow(training)
training <- training[, na_counts < na_threshold]

# Keep only relevant features
testing <- testing[, intersect(names(training), names(testing)), drop = FALSE]

# Convert target variable to factor
training$classe <- as.factor(training$classe)
```

---

## Model Training & Selection

A **Random Forest** model is trained with 70% of the dataset for training and 30% for validation. We evaluate its performance using cross-validation.

```{r}
set.seed(456)
trainIndex <- createDataPartition(training$classe, p = 0.7, list = FALSE)
trainData <- training[trainIndex, ]
validData <- training[-trainIndex, ]

# Train Random Forest model
rf_model <- randomForest(classe ~ ., data = trainData, ntree = 150)

# Save trained model
saveRDS(rf_model, "rf_model.rds")
```

---

## Model Evaluation

The model is evaluated using a confusion matrix and feature importance.

```{r}
rf_predictions <- predict(rf_model, validData)
conf_matrix <- confusionMatrix(rf_predictions, validData$classe)
print(conf_matrix)

# Improved Feature Importance Visualization
importance_df <- as.data.frame(importance(rf_model))
importance_df$Feature <- rownames(importance_df)
importance_df <- importance_df %>% arrange(desc(MeanDecreaseGini))

# Plot using ggplot2
ggplot(importance_df, aes(x = reorder(Feature, MeanDecreaseGini), y = MeanDecreaseGini)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Feature Importance", x = "Features", y = "Importance Score") +
  theme_minimal()
```

---

## Final Predictions

Predictions on new test cases are generated, and results are saved directly in the working directory.

```{r}
if (nrow(testing) > 0) {
  final_predictions <- predict(rf_model, testing)

  # Save predictions
  for (i in 1:length(final_predictions)) {
    filename <- paste0("prediction_result_", i, ".txt")
    write.table(final_predictions[i], file = filename, quote = FALSE, row.names = FALSE, col.names = FALSE)
  }
}
```

---

## Conclusion

This report demonstrates the classification of exercise types using machine learning. The **Random Forest model** provides high accuracy, and predictions on new test cases are generated successfully.
